A state space model reveals the rich and varied dynamics of replay

### 2020-07-15 Article.tex
Update on Overleaf.

Our results indicate that most events with coherent spatial content are not continuous; that is, trajectories that move many times the running speed of the animal. We therefore examined these events in more detail. We first noted that most of the SWRs (16454 of 23382 classified SWRs or 70\%) were well described by a single dynamic. Surprisingly, the most common SWRs were categorized exclusively as stationary-continuous mixtures (9860 of 23382 classified SWRs or 42\%, Figure \ref{5}A). These events contain representations that move at rates more consistent with real-world movement (Fig. 1G), but were slightly shorter, on average, than events with continuous dynamics (median duration: stationary-continuous-mixture 72 ms vs. continuous 95 ms, Figure \ref{5}B). Nonetheless, both these slow-moving events and continuous events represented locations that were some distance away from the animal (mean trajectory distance from animal's position: stationary-continuous-mixture 51 cm vs. continuous 42 cm, Figure \ref{5}C). This indicates that the trajectories of these SWRs, like ones that have faster continuous dynamics, do not represent the animal's actual position.

@@

\section*{Discussion}
Here we developed a state space model that identifies and quantifies generative activity in terms of a mixture of three movement dynamics: stationary, continuous, and fragmented. We showed that this model is interpretable; it has well-defined parameters and allows us to explain a large fraction of SWRs in terms of these dynamics---far greater than had been explained before. We showed that this model is flexible; it can be used with sorted or clusterless spikes, it can be used with 1D or 2D positions, and---because it uses a mixture of dynamics---it allowed us to confidently discover SWRs with non-constant speed and non-linear trajectories that would be difficult to characterize with the typical Bayesian decoder. Finally, we showed that this model is robust; it was relatively insensitive to transition parameters and depended strongly on the hippocampal place representation structure.

Using this model, we found that most SWRs contained spatially coherent content. Importantly, this content was not predominantly trajectories that occurred at many times the speed of the animal, but trajectories that were much closer to the animal's actual movement speed. Many of these events did not simply represent the animal's current location, but other locations on track.

Our results challenge the notion that hippocampal replays are necessarily many times the speed of the animal. Are stationary events during SWR that represent the animal's position different from stationary events that represent non-local positions

>>>>

Our results indicate that most events with coherent spatial content are not continuous; that is, trajectories that move many times the running speed of the animal. We therefore examined these events in more detail. We first noted that most of the SWRs (16454 of 23382 classified SWRs or 70\%) were well described by a single dynamic. Surprisingly, the most common SWRs were categorized exclusively as stationary-continuous mixtures (9860 of 23382 classified SWRs or 42\%, Figure \ref{5}A). These events contain representations that move at rates more consistent with real-world movement (Fig. 1G), but were slightly shorter, on average, than events with continuous dynamics (median duration: stationary-continuous-mixture 72 ms vs. continuous 95 ms, Figure \ref{5}B). Nonetheless, both these slow-moving events and continuous events represented locations that were some distance away from the animal (mean trajectory distance from animal's position: stationary-continuous-mixture 51 cm vs. continuous 42 cm, Figure \ref{5}C). This indicates that the trajectories of these SWRs, like ones that have faster continuous dynamics, do not represent the animal's actual position and cannot be explained by erroneously detected SWRs.

@@

\section*{Discussion}
Here we developed a state space model that identifies and quantifies generative activity in terms of a mixture of three movement dynamics: stationary, continuous, and fragmented. We showed that this model is interpretable; it has well-defined parameters and allows us to explain a large fraction of SWRs in terms of these dynamics---far greater than had been explained before. We showed that this model is flexible; it can be used with sorted or clusterless spikes, it can be used with 1D or 2D positions, and---because it uses a mixture of dynamics---it allowed us to confidently discover SWRs with non-constant speed and non-linear trajectories that would not be found with the typical Bayesian decoder. Finally, we showed that this model is robust; it was relatively insensitive to misspecified transition parameters between dynamics and depended strongly on the hippocampal place representation structure.

Using this model, we found that most SWRs contained spatially coherent content. Importantly, this content was not predominantly trajectories that occurred at many times the speed of the animal, but trajectories that were much closer to the animal's actual movement speed. Many of these events did not simply represent the animal's current location, but other locations on track. This result challenges the long-held notion that hippocampal replay trajectories necessarily proceed at many times the speed of the animal \cite{NadasdyReplayTimeCompression1999}. Are stationary events during SWR that represent the animal's position different from stationary events that represent non-local positions
